---
title: "LFulton_Assignment7_JIden"
author: "Josh Iden"
date: "2022-10-09"
output: html_document
---

### HW7

#### 1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi's. Find the distribution of Y

Let $k^n$ be the combinations for $Xi$ for integers $1$ to $k$   

Let $j$ be the combinations for minimum $k^n$

$Y = 1 \le j \le k, \frac{(k - j + 1)^n - (k - j)^n}{k^n}$

#### 2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.)

##### a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

The expected value is given: $E(X) = 10$
The probability is given: $p = \frac{1}{10} = 0.1$
Standard deviation = $\frac{\sqrt{1-p}}{p}$

```{r}
p <- 0.1 #probability
std <- round((1-p)**0.5/p,2)
print(paste("Standard Deviation:",std))
```


The probability function for a geometric distribution:
$P(X = x) = (1 - p)^{x-1} * p =>$

With X = 8 (years):
$P(X = 8) = (1 - p)^{8-1} * p$

in R:

```{r}
x <- 8 #number of years
pgeom(x, p) 
```

= the probability the machine will fail after 8 years (will not fail in first 8 years) as a geometric distribution.

#### b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

```{r}
pexp(x, p)
```

The standard deviation in an exponential distribution is equal to the expected value: 10.

#### c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

```{r}
pbinom(0, x, p, lower.tail=FALSE)
```

Computing the expected value and standard deviation:

```{r}
exp <- x * p
var <- x * p * (1-p)
std <- round(sqrt(var),2)

print(paste("Expected value:",exp))
print(paste("Standard deviation:",std))
```

#### d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

```{r}
lambda <- 10 #expected value
st_d <- round(sqrt(lambda),2) #standard deviation
prob <- round(ppois(x, lambda, lower.tail=FALSE),2)

print(paste("Expected value:",lambda))
print(paste("Standard deviation",st_d))
print(paste("Probability of failure after 8 years:",prob))
```

