---
title: "HW2"
author: "Josh Iden"
date: "2022-09-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA 605: Assignment 2

### Problem Set #1

*1) Show that* $A^tA ≠ AA^t$ *in general. Proof and demonstration.*

We know from its definition that matrix multiplication is **not commutative**, so, generally speaking $A^tA ≠ AA^t$.

Example:

```{r}
# define a 2x2 matrix
A <- matrix(seq(1,4),2)
print(A)
```

```{r}
# define the transpose 
T <- t(A)
print(T)
```

```{r}
# check for equality
T%*%A == A%*%T
```

*2) For a special type of square matrix, we get* $A^tA ≠ AA^t$. *Under what conditions is this true?*

Square matrices **are commutative** when they are **symmetrical**, meaning $A^t = A$.

Example:

```{r}
# define a symmetrical matrix
A <- matrix(c(1,2,2,4),nrow=2)
A
```

```{r}
# take the transpose of the matrix
T <- t(A)
T
```

```{r}
# check for equality
T%*%A == A%*%T
```

### Problem Set #2

*Write an R function to factorize a square matrix A into LU or LDU*

This rudimentary function takes a 2x2 matrix and returns the lower and upper triangular matrices:

```{r}
library(matlib)
library(pracma)

# define a square matrix
A <- matrix(c(2,6,1,8),2)

## define function LUD 2x2
LUD2 <- function(A) {
  # define row coefficient: B
  B <- A[2,1] / A[1,1]
  # define upper triangular matrix: U
  U <- A
  U[2.1] <- A[2,1] - B * A[1,1]
  U[2,2] <- A[2,2] - B * A[1,2]
  # define lower triangular matrix: L
  L <- diag(dim(A)[1])
  L[2,1] <- -B
  output <- list("Lower"=L, "Upper"=U)
  return(output)
}

LUD2(A)
```

Building on the above, this function takes larger matrices and returns the lower and upper triangular matrices.

```{r}
# define a 3x3 matrix
D <- matrix(c(1,2,3,4,5,6,7,8,9), 3)

## define function LUD3
LUD3 <- function(A) {
  # check dimension
  if (dim(A)[1] != dim(A)[2]) stop ("not a square matrix")
  # initialze uppper triangle: U
  U <- A
  # initialize lower triangle: L
  L <- diag(dim(A)[1])
  
  # iterate through rows
  for (i in 1:nrow(A)) {
    for (j in 2:nrow(A)) {
      if (j > i) {
        # define row val: x
        x <- U[i,]
        # define row coefficient: B
        B <- U[j,i] / x[i]
        U[j,] <- U[j,] - (B * x)
        L[j,i] <- B
      }
    }
  }
  
  output <- list("Lower"=L, "Upper"=U)
  return(output)
}

LUD3(D)
```
